---
title: "Lab 5"
author: "An Bui"
date: "7 February 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r libraries and data, message = FALSE}
#### libraries ####
library(tidyverse)
library(tseries)
library(forecast)

#### data ####
# US residential energy consumption values (Jan 1973 - Oct 2017, trillion BTU)
energy <- read_csv("energy.csv")
```

First, create 'ts' time series data using information in the res_total column.
```{r time series, message = FALSE}
#### set data as time series ####
res_ts <- ts(energy$res_total,
             # monthly data
             frequency = 12,
             # start January 1973
             start = c(1973, 1))
res_ts

#### plot data against time ####
plot(res_ts)
# looks like there's an increasing trend
# looks additive: highest highs and lowest lows within season remain relatively constant
```

Now decompose the data.
```{r decompose, message = FALSE}
#### decompose function ####
res_dc <- decompose(res_ts)

#### plot ####
plot(res_dc)
# increasing trend
# seasonality (on scale of 100s, so seasonality is big contributor to patterns observed)
# no pattern to residuals
```

There are other ways to visualize the data.
```{r other visualizations, message = FALSE}
#### plot months ####
monthplot(res_ts)
# breaks down within month how values are changing
# mean value for each month over the years

#### season plot ####
ggseasonplot(res_ts)
# lower orange values are in the 1970s, purple pink values are recent
# every year has its own series, tells the same story in a different way

ggseasonplot(res_ts, polar = TRUE)
# plot in a circle
```

##### 1. Simple moving averages.
Use `ma()` in `forecast` package.
```{r moving averages, message = FALSE}
sma_res <- ma(res_ts,
              # window size
              order = 5)
# first 2 and last 2 observations will get cut because of the number of observations/5

plot(res_ts)
lines(sma_res, col = "red")
# run together to get data and moving average
# looks smooth

sma_res_12 <- ma(res_ts,
              # window size
              order = 12)
plot(sma_res_12)
# looks even smoother

sma_res_50 <- ma(res_ts,
              # window size
              order = 50)
plot(sma_res_50)
# even smoother
```
Clearly, changing windows changes the way the data appear.

##### 2. Autocorrelation
```{r acf, message = FALSE}
res_acf <- acf(res_ts)
# big peaks every 12 months, which means that things are highly correlated with observations 12 months prior
# makes sense for energy usage data

ggtsdisplay(res_ts)
# original data, acf, and pacf
```
Seasonality is important!

##### 3. Holt-Winters exponential smoothing
```{r holt-winters, message = FALSE}
#### holt-winters ####
res_hw <- HoltWinters(res_ts)
res_hw
plot(res_hw)
# shows original data in black, model predictions in red

#### forecasting ####
res_forecast <- forecast(res_hw,
                         # how many increments to forecast by (60/5 = 12 months)
                         h = 60)
plot(res_forecast)
# slightly downward trend

#### check distribution of residuals ####
hist(res_forecast$residuals)
# look normally distributed
```

##### 4. Autoregressive integrated moving average (ARIMA)
```{r arima, message = FALSE}
#### find pdq ####
res_pdq <- auto.arima(res_ts)
res_pdq
# pdq (nonseasonal): 1, 0, 2
# pdq (seasonal): 0, 1, 1

#### fit ARIMA model using pdq inputs ####
res_arima <- arima(res_ts,
                   # give non-seasonal pdq
                   order = c(1, 0, 2),
                   # give seasonal pdq as list
                   seasonal = list(order = c(0, 1, 1)))
res_arima

#### look at residuals ####
# set parameters for graphics: populate graphics in 1 row with 2 columns to get graphs side by side
par(mfrow = c(1, 2))
hist(res_arima$residuals)
qqnorm(res_arima$residuals)
# looks normally distributed ish - not skewed in either direction
```

##### 5. Forecast with ARIMA outputs.
```{r forecast, message = FALSE}
#### forecast ####
forecast_res <- forecast(res_arima,
                         # forecast to 72 months
                         h = 72)
par(mfrow = c(1, 1))
plot(forecast_res)

#### graph with ggplot ####
res_df <- data.frame(forecast_res)
# row names are months and years
# point forecast, lower and upper bounds for 80th and 95th % confidence interval

# make a sequence of numbers for months
month_seq <- seq(1, 72)
# bind to data frame
res_df_2 <- data.frame(month_seq, res_df)

ggplot(res_df_2, aes(x = month_seq,
                     y = Point.Forecast)) +
  geom_line()
```












